---
title: "Karin"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(plotly)
library(dplyr)

data(bank)
bank <-read.table("http://www.trutschnig.net/Datensatz.txt",head=TRUE)

data(RTR2015)
address <- url("http://www.trutschnig.net/RTR2015.RData")
load(address)
head(RTR2015)

```



## Overview

- Basics
- Intermediate
- Application Time Series
- Statistics
- Mean Shift Clustering
.......


## Useful Links 

https://plot.ly/r/
  
https://plotly-book.cpsievert.me/
  
https://help.plot.ly/tutorials/
  
  
## Let?s get started:

## Installation: 
```
install.packages("plotly")
library(plotly)
library(dplyr)
```


#Readin Bankomatdatensatz
```
bank <-read.table("http://www.trutschnig.net/Datensatz.txt",head=TRUE)
summary(bank)
```


#Readin RTR data (see https://www.netztest.at/de/Test for Background)
```
address <- url("http://www.trutschnig.net/RTR2015.RData")
load(address)
head(RTR2015)
```


## BASISC

The general structure of a plotly command looks like this:

```
plot_ly(data, x = ~X, y = ~Y, ...)
```


If you do not specify the type of plot you want, plotly will choose the one that suits the given data best.

```
Galaxy = filter(RTR2015, device == "Galaxy Note 10.1") 

plot_ly(Galaxy, x = ~rtr_speed_ul, y = ~rtr_speed_dl, name = "default")
```

  
```{r}
Galaxy = filter(RTR2015, device == "Galaxy Note 10.1") 

plot_ly(Galaxy, x = ~rtr_speed_ul, y = ~rtr_speed_dl, name = "default")
```

BUT: 

```
plot_ly(Galaxy, x = ~op_name, y = ~rtr_speed_dl, name = "default")
```
```{R}
plot_ly(Galaxy, x = ~op_name, y = ~rtr_speed_dl, name = "default")
```


Scatter:

The scatterplot is useful for visualizing the correlation between variables.

```
plot_ly(Galaxy, x = ~ rtr_speed_ul, y = ~rtr_speed_dl)%>% 
  add_markers(alpha = 0.9, name = "alpha")
```
```{R}  
plot_ly(Galaxy, x = ~ rtr_speed_ul, y = ~rtr_speed_dl)%>% 
  add_markers(alpha = 0.9, name = "alpha")
```

```
plot_ly(Galaxy, x = ~ rtr_speed_ul, y = ~rtr_speed_dl)%>% 
  add_markers(symbol = I(2), name = "hollow")
```
```{R}  
plot_ly(Galaxy, x = ~ rtr_speed_ul, y = ~rtr_speed_dl)%>% 
  add_markers(symbol = I(2), name = "hollow")
```

```
subplot(
  plot_ly(Galaxy, x = ~rtr_speed_ul, y = ~rtr_speed_dl, name = "default"),
  plot_ly(Galaxy, x = ~rtr_speed_ul, y = ~rtr_speed_dl) %>% 
    add_markers(alpha = 0.2, name = "alpha"),
  plot_ly(Galaxy, x = ~rtr_speed_ul, y = ~rtr_speed_dl) %>% 
    add_markers(symbol = I(1), name = "hollow")
)
```
```{R}

subplot(
  plot_ly(Galaxy, x = ~rtr_speed_ul, y = ~rtr_speed_dl, name = "default"),
  plot_ly(Galaxy, x = ~rtr_speed_ul, y = ~rtr_speed_dl) %>% 
    add_markers(alpha = 0.2, name = "alpha"),
  plot_ly(Galaxy, x = ~rtr_speed_ul, y = ~rtr_speed_dl) %>% 
    add_markers(symbol = I(1), name = "hollow")
)
```

```
subplot(
  plot_ly(Galaxy, x = ~rtr_speed_ul, y = ~rtr_speed_dl, name = "default"),
  plot_ly(Galaxy, x = ~rtr_speed_ul, y = ~rtr_speed_dl) %>% 
    add_markers(alpha = 0.5, name = "alpha"),
  plot_ly(Galaxy, x = ~rtr_speed_ul, y = ~rtr_speed_dl) %>% 
    add_markers(symbol = I(3), name = "hollow")
)
```

```{R}

subplot(
  plot_ly(Galaxy, x = ~rtr_speed_ul, y = ~rtr_speed_dl, name = "default"),
  plot_ly(Galaxy, x = ~rtr_speed_ul, y = ~rtr_speed_dl) %>% 
    add_markers(alpha = 0.5, name = "alpha"),
  plot_ly(Galaxy, x = ~rtr_speed_ul, y = ~rtr_speed_dl) %>% 
    add_markers(symbol = I(3), name = "hollow")
)
```



We can map additional information by using colors:

```
plot_ly(data = Galaxy, x = ~rtr_speed_ul, y = ~rtr_speed_dl, color = ~op_name)
```

```{R}
plot_ly(data = Galaxy, x = ~rtr_speed_ul, y = ~rtr_speed_dl, color = ~op_name)
```
```
plot_ly(data = Galaxy, x = ~rtr_speed_ul, y = ~rtr_speed_dl, color = ~op_name, colors = c("red","green","blue"))
```
```{R}
plot_ly(data = Galaxy, x = ~rtr_speed_ul, y = ~rtr_speed_dl, color = ~op_name, colors = c("red","green","blue"))
```

Another option is add_lines

```
plot_ly(Galaxy, x = ~longitude, y = ~rtr_speed_dl) %>%
  add_lines(color = ~op_name)
```
```{R}  
plot_ly(Galaxy, x = ~longitude, y = ~rtr_speed_dl) %>%
  add_lines(color = ~op_name)
```

## Exercise:
Choose some other devices and visualize the correlation between download-Speed and Ping. 


## Histograms and Bars

A histogram is an accurate representation of the distribution of numerical data. It differs from a bar graph, in the sense that a bar graph relates two variables, but a histogram relates only one. 

```
plot_ly(RTR2015, x = ~rtr_speed_dl, color = ~device_platform) %>% add_histogram(name = "plotly.js")
```
```{R}
plot_ly(RTR2015, x = ~rtr_speed_dl, color = ~device_platform) %>% add_histogram(name = "plotly.js")
```

R has diverse facilities for estimating the optimal number of bins in a histogram that we can easily leverage.

```
p1 <- plot_ly(RTR2015, x = ~rtr_speed_dl) %>% add_histogram(name = "plotly.js")
speed_hist <- function(method = "FD") {
  h <- hist(RTR2015$rtr_speed_dl, breaks = method, plot = FALSE)
  plot_ly(x = h$mids, y = h$counts) %>% add_bars(name = method)
}
```
```{R}
p1 <- plot_ly(RTR2015, x = ~rtr_speed_dl) %>% add_histogram(name = "plotly.js")
speed_hist <- function(method = "FD") {
  h <- hist(RTR2015$rtr_speed_dl, breaks = method, plot = FALSE)
  plot_ly(x = h$mids, y = h$counts) %>% add_bars(name = method)
}
```


```
subplot(
  p1, speed_hist(), speed_hist("Sturges"),  speed_hist("Scott"),
  nrows = 4, shareX = TRUE
)
```

```{R}
subplot(
  p1, speed_hist(), speed_hist("Sturges"),  speed_hist("Scott"),
  nrows = 4, shareX = TRUE
)
```
Other Examples

```
plot_ly(RTR2015, x = ~op_name) %>% add_histogram()
```
```{R}
plot_ly(RTR2015, x = ~op_name) %>% add_histogram()
```


```
plot_ly(RTR2015, x = ~op_name, color = ~device_platform) %>% add_histogram()
```
```{R}
plot_ly(RTR2015, x = ~op_name, color = ~device_platform) %>% add_histogram()
```
```
plot_ly(RTR2015, x = ~op_name, y =~rtr_speed_dl, color = ~device_platform) %>% add_bars()
```
```{R}
plot_ly(RTR2015, x = ~op_name, y =~rtr_speed_dl, color = ~device_platform) %>% add_bars()
```


```
plot_ly(RTR2015, x = ~device_has_lte, y= ~device, color = ~op_name) %>%
  add_bars() %>%
  layout(barmode = "stack")
```

```{R}
plot_ly(RTR2015, x = ~device_has_lte, y= ~device, color = ~op_name) %>%
  add_bars() %>%
  layout(barmode = "stack")
```

Exercise:

Starting from those data where op_name = "Apple"

-create a histogram showing the distribution of rtr_ping
-create a barplot of nw_cat and rtr_Speed_dl 
-create stacked bars which in addition also contain the device platform

## STATISTIC 

## Boxplots

A boxplot is a way to show the spread and centers of a data set. Measures of spread include the interquartile range and the mean of the data set. Measures of center include the mean or average and median (the middle of a data set).

```
plot_ly(Galaxy, y = ~rtr_speed_ul, type = "box", name ="Upload") %>%
  add_trace(y = ~rtr_speed_dl, name = "Download")
```
```{R}
plot_ly(Galaxy, y = ~rtr_speed_ul, type = "box", name ="Upload") %>%
  add_trace(y = ~rtr_speed_dl, name = "Download")
```


```
plot_ly(Galaxy, x = ~rtr_speed_ul, type = "box", name ="Upload") %>%
  add_trace(x = ~rtr_speed_dl, name = "Download")
```
```{R}
plot_ly(Galaxy, x = ~rtr_speed_ul, type = "box", name ="Upload") %>%
  add_trace(x = ~rtr_speed_dl, name = "Download")
```



```
p <- plot_ly(Galaxy, y = ~rtr_speed_dl, color = I("blue"), 
             alpha = 0.1, boxpoints = "suspectedoutliers")
p1 <- p %>% add_boxplot(x = "Overall")
p2 <- p %>% add_boxplot(x = ~op_name)
subplot(
  p1, p2, shareY = TRUE,
  widths = c(0.2, 0.8), margin = 0
) %>% hide_legend()
```
```{R}
p <- plot_ly(Galaxy, y = ~rtr_speed_dl, color = I("blue"), 
             alpha = 0.1, boxpoints = "suspectedoutliers")
p1 <- p %>% add_boxplot(x = "Overall")
p2 <- p %>% add_boxplot(x = ~op_name)
subplot(
  p1, p2, shareY = TRUE,
  widths = c(0.2, 0.8), margin = 0
) %>% hide_legend()
```


```
d = filter(RTR2015, nw_cat %in% c("4G", "3G", "2G"))
plot_ly(d, x = ~rtr_speed_dl, y = ~interaction(nw_cat, device_has_lte)) %>%
  add_boxplot(color = ~nw_cat) %>%
  layout(yaxis = list(title = ""), margin = list(l = 100))
```

```{R}
d = filter(RTR2015, nw_cat %in% c("4G", "3G", "2G"))
plot_ly(d, x = ~rtr_speed_dl, y = ~interaction(nw_cat, device_has_lte)) %>%
  add_boxplot(color = ~nw_cat) %>%
  layout(yaxis = list(title = ""), margin = list(l = 100))
```


## Exercise: 
Use Boxplots to compare the Upload-Speed of devices that have LTE and the ones that don?t 

## Heatmaps

```
plot_ly(RTR2015, x = ~op_name, y = ~nw_cat, z = ~rtr_speed_dl, type = "heatmap")
```
```{R}
plot_ly(RTR2015, x = ~op_name, y = ~nw_cat, z = ~rtr_speed_dl, type = "heatmap")
```


```
plot_ly(RTR2015, x = ~mymd, y = ~device_platform, z = ~rtr_speed_dl, type = "heatmap")
```

```{R}
plot_ly(RTR2015, x = ~mymd, y = ~device_platform, z = ~rtr_speed_dl, type = "heatmap")
```

```
plot_ly(RTR2015, x = ~mymd, y = ~op_name, z = ~rtr_speed_dl, type = "heatmap")
```

```{R}
plot_ly(RTR2015, x = ~mymd, y = ~op_name, z = ~rtr_speed_dl, type = "heatmap")
```

```
plot_ly(RTR2015, x = ~mymd, y = ~op_name, z = ~rtr_speed_dl, type = "heatmap", colorscale="Greys")
```
```{R}
plot_ly(RTR2015, x = ~mymd, y = ~op_name, z = ~rtr_speed_dl, type = "heatmap", colorscale="Greys")
```


```
plot_ly(RTR2015, x = ~mymd, y = ~op_name, z = ~rtr_speed_dl,colors = colorRamp(c("red", "green")), type = "heatmap")
```
```{R}
plot_ly(RTR2015, x = ~mymd, y = ~op_name, z = ~rtr_speed_dl,colors = colorRamp(c("red", "green")), type = "heatmap")
```


```
plot_ly(RTR2015, x = ~op_name, y = ~nw_cat, z = ~rtr_speed_dl, colors = colorRamp(c("red", "yellow")), type = "heatmap")
```


```{R}
plot_ly(RTR2015, x = ~op_name, y = ~nw_cat, z = ~rtr_speed_dl, colors = colorRamp(c("red", "yellow")), type = "heatmap")
```

## Exercise: 
  Create a heatmap visualising the connection of nw_cat, mymd and rtr_speed_ul

## 2DHistograms 

2D histograms require x/y, but in contrast to heatmaps, z is optional

```
plot_ly(RTR2015, x = ~op_name, y = ~nw_cat) %>%
 add_histogram2d()
```
```{R}
plot_ly(RTR2015, x = ~op_name, y = ~nw_cat) %>%
 add_histogram2d()
``` 
 
```
plot_ly(RTR2015, x = ~op_name, y = ~rtr_speed_dl) %>%
  add_histogram2d()
```
```{R}
plot_ly(RTR2015, x = ~op_name, y = ~rtr_speed_dl) %>%
  add_histogram2d()
```

